---
title: "Project Proposal"
author: "Benjamin Leidig, Monte Thomas, Harmony Pham"
date: "2025-03-26"
output: pdf_document
---

### Section: GR




### Research Questions
* How can we forecast daily hotel total revenue using historical revenue data?
* Are there seasonal fluctuations in daily total hotel revenue?
* Do any external events (sport events, family weekends, etc.) effect daily hotel total revenue?




### Potential Sources
* (https://medium.com/@chenycy/predict-hotel-demands-leveraging-time-series-forecasting-techniques-62e25606f273)
* (https://pure.psu.edu/en/publications/forecasting-hotel-occupancy-rates-with-time-series-models-an-empi)




```{r setup, include=FALSE, root.dir = "C:/Users/dacn/Desktop/429-project"}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tseries)
library(astsa)
```



\newpage
### Dataset Cleaning

```{r}
# Create a function to clean data
clean_hotel_data <- function(file_path, year){
  read.csv(file_path) %>%
    pivot_longer(cols = -X, names_to = "Date", values_to = "Value") %>%
    pivot_wider(names_from = "X", values_from = "Value") %>%
    mutate(
      Date = str_remove(Date, "^[A-Za-z]+") %>% 
             paste0(".", year) %>%  
             mdy(),
      Weekday = wday(Date, label = TRUE)
    ) %>%
    select(Date, Weekday, Occupancy, ADR, TOTAL_REVENUE)
}
```

```{r}
# Apply function to data1 (year 2024)
data1 <- clean_hotel_data("hoteldata24.csv", "2024")

# Apply function to data2 (year 2025)
data2 <- clean_hotel_data("hoteldata25.csv", "2025")

# Combine datasets
hotel_data <- bind_rows(data1, data2)
str(hotel_data)
head(hotel_data)

# Saving the new dataset as a CSV
write.csv(hotel_data, "cleaned_hotel_data.csv", row.names = FALSE)
```



\newpage
## Visualizations

```{r}
summary(hotel_data)
```

\newpage
### Occupancy Rate Visualizations

```{r}
# Plot Occupancy over Time
ggplot(hotel_data, aes(x = Date, y = Occupancy)) +
  geom_line(color = "black", linewidth = 0.5) +
  geom_point(color = "black", size = 1) +
  geom_smooth(method = "lm", color = "red", se = FALSE, linewidth = 1, linetype = 2) +
  labs(title = "Hotel Occupancy Over Time",
       x = "Date",
       y = "Occupancy") +
  theme_minimal()
```
```{r}
# Plot Occupancy distribution
ggplot(hotel_data, aes(x = Occupancy)) +
  geom_histogram(bins = 10, fill = "gray", color = "black") +
  labs(title = "Distribution of Occupancy",
       x = "Occupancy",
       y = "Frequency") +
  theme_minimal()
```
```{r}
acf(hotel_data$Occupancy, main = 'Occupancy')
adf.test(hotel_data$Occupancy); kpss.test(hotel_data$Occupancy, null = 'Trend')
```


\newpage
### ADR Visualizations

```{r}
# Plot ADR over Time
ggplot(hotel_data, aes(x = Date, y = ADR)) +
  geom_line(color = "black", linewidth = 0.5) +
  geom_point(color = "black", size = 1) +
  geom_smooth(method = "lm", color = "red", se = FALSE, linewidth = 1, linetype = 2) +
  labs(title = "Hotel ADR Over Time",
       x = "Date",
       y = "ADR") +
  theme_minimal()
```
```{r}
# Plot ADR distribution
ggplot(hotel_data, aes(x = ADR)) +
  geom_histogram(bins = 10, fill = "grey", color = "black") +
  labs(title = "Distribution of ADR",
       x = "ADR",
       y = "Frequency") +
  theme_minimal()
```
```{r}
acf(hotel_data$ADR, main = 'ADR')
adf.test(hotel_data$ADR); kpss.test(hotel_data$ADR, null = 'Trend')
```

\newpage
### Total Revenue Visualizations

```{r}
# Plot TOTAL_REVENUE over Time
ggplot(hotel_data, aes(x = Date, y = TOTAL_REVENUE)) +
  geom_line(color = "black", linewidth = 0.5) +
  geom_point(color = "black", size = 1) +
  geom_smooth(method = "lm", color = "red", se = FALSE, linewidth = 1, linetype = 2) +
  labs(title = "Hotel TOTAL_REVENUE Over Time",
       x = "Date",
       y = "TOTAL_REVENUE") +
  theme_minimal()
adf.test(hotel_data$TOTAL_REVENUE); kpss.test(hotel_data$TOTAL_REVENUE, null = 'Trend')
```

```{r}
# Plot TOTAL_REVENUE distribution
ggplot(hotel_data, aes(x = TOTAL_REVENUE)) +
  geom_histogram(bins = 10, fill = "grey", color = "black") +
  labs(title = "Distribution of TOTAL_REVENUE",
       x = "TOTAL_REVENUE",
       y = "Frequency") +
  theme_minimal()
```
```{r}
acf(hotel_data$TOTAL_REVENUE, main = 'Total Revenue')
adf.test(hotel_data$TOTAL_REVENUE); kpss.test(hotel_data$TOTAL_REVENUE, null = 'Trend')
```


\newpage
### Dataset Description

For this project, we are using real data collected from the I Hotel & Illinois Conference Center. The dataset--stored in the dataframe object, `hotel_data`--consists of five variables: `Date` (YYYY-MM-DD), `Weekday` (Sun, Mon, Tue, Wed, Thu, Fri, Sat), `Occupancy` (a percentage represented as a decimal), `ADR` (Average Daily Rate; (room revenue)/(rooms sold); average daily revenue in USD earned per occupant), and `TOTAL_REVENUE` (total daily revenue in USD). Each observations represents an individual day, between 2024/03/28 and 2025/03/27.

Occupancy rates tends to be skewed to the left, with a mode of approximately 0.9. Looking at the time series plot of occupancy rates, there appears to be a non-constant mean function that is decreasing, although there is also a potential seasonal component as well. There appears to be a non-constant variance that increases with time. An ADF test yields a p-value of less than 0.01, meaning that the time series could be stationary. However, a KPSS test (with $H_0:\text{trend stationarity}$) also yields a p-value of less than 0.01, which means the time series isn't trend-stationary, either. Although these test are useful, we conclude via visual inspection that the series is neither stationary nor trend-stationary.

ADR tends to be skewed to the right, with a mode of approximately 150. Looking at the time series plot of ADR, there appears to be a constant mean function, although there is also a strong seasonal component. There appears to be a non-constant variance, with variance increasing during peak seasons and decreasing during the depressions. An ADF test yields a p-value of less than 0.01, meaning that the time series could be stationary. However, a KPSS test (with $H_0:\text{trend stationarity}$) yields a p-value of 0.02373, which, if using a significance level of 0.05, means the time series isn't trend stationary. Although these test are useful, we conclude via visual inspection that the series is neither stationary nor trend-stationary.

Total revenue tends to be skewed to the right, with a mode of approximately 150. Looking at the time series plot of total revenue, there appears to be a non-constant mean function. There is also not a prominent seasonal component. There appears to be a non-constant variance, with variance increasing through April to May and September to November. An ADF test yields a p-value of less than 0.01, meaning that the time series could be stationary. However, a KPSS test (with $H_0:\text{trend stationarity}$) yields a p-value of 0.01311, which, if using a significance level of 0.05, means the time series isn't trend stationary. Although these test are useful, we conclude via visual inspection that the series is neither stationary nor trend-stationary.

According to the sample ACF plots, there is a strong seasonal component in all of the time series. In particular, starting at lag 0 ($h=0$), the sample ACF peaks at increments of 7 (i.e. at $h=7$, $h=14$, $h=21$, etc.). This implies that all three variables have an association with the day of the week.